\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-10-26 syntax.tex
Renaming Typing Rules, Part 1 Use this git change to rename all the legacy name from the new rule name

% Example : how judgemental equality is used for the universe
We review the concept of ``universe'' and at the same time showing how equality is representing semantic. In dependent type setting, function now can return type, of which the type is ``type of type'', i.e., \textit{the universe} $\cU$. We follow the style of \textit{type universe à la Tarski}~\cite{hofmann1997syntax} and
distinguish types from their \textit{codes or names} as showin in \ruleref{Code/Get Type}: given a \textit{code} (term) $\goodTerm{\Gamma}{T}{\cU}$ we can retrieve the corresponding type
$\goodType{\Gamma}{El~T}{}$, and vice-versa.\YZ{Why is it wanted to distinguish types from their codes?}

@@

\label{fig:rules:well-typed-ctx}
\judgebox{\goodCtx{\Gm}{i}}
$$ 
\Rule[name=Empty Context]{}{\goodCtx{\cdot}{0}} 
\quad
\Rule[name=Context Extension]
{\goodCtx{\Gm}{i} \quad \goodType{\Gm}{A}{j}}
{\goodCtx{\Gm, A}{i \cup j}}  
$$

\judgebox{ \goodType{\Gm}{T}{i} }
$$
\Rule[name=Type Universe]
{}
{\goodType{\Gm}{\cU}{j + 1}}
\quad 
\Rule[name=Boolean]
{}
{\goodType{\Gm}{\cB}{0}}
\quad 
\Rule[name=Bottom]
{}
{\goodType{\Gm}{\bot}{0}}
\quad 
\Rule[name=Function]
{\goodType{\Gm}{A}{j} 
  \quad \goodType{\Gm, A}{B}{k}}
{\goodType{\Gm}{\TyPi{A}{B}}{j \cup k}}
$$

@@

$$
\Rule[name=Bool]
{}
{\goodTerm{\Gm}{\true, \false}{\cB}}
\quad
\Rule[]
{\goodSub{\Gm}{\gm}{\Delta}}
{\goodType{\Gm}{ \sub{(\El{T})}{\gm} \equiv \El{(\sub{T}{\gm})} }{j} \quad
 \goodTerm{\Gm}{ \sub{\true}{\gm} \equiv \true}{\cB} \quad 
 \goodTerm{\Gm}{ \sub{\false}{\gm} \equiv \false}{\cB} 
}
$$
\judgebox{\goodSub{\Gm}{\gm}{\Delta}}
$$
\Rule[name=Ept Subst]
{}{\goodSub{\Gm}{\epsilon}{\cdot}}
\quad
\Rule[]
{}{\goodSub{\Gm}{"id"}{\Gm}}
\quad
\Rule[name=Sub Comp]{
  \goodSub{\Delta}{\delta}{\Theta}
  \quad \goodSub{\Gm}{\gm}{\Delta} 
}{\goodSub{\Gm}{ \SubstComp{\delta}{\gm}}{\Theta}}
\Rule[name=Sub Extend]
{\goodSub{\Gm}{\gm}{\Delta} \quad \goodTerm{\Gm}{t}{ \sub{A}{\gm}}}
{\goodSub{\Gm}{(\gm, t)}{(\Delta, A)}}
$$

@@

A term $\goodTerm{\Gamma}{t}{A}$ can be understood as one term $t$ of
type $A$, with a bunch of holes (of type $\Gamma$) inside the term $t$.
A \textbf{simultaneous substitution} $\goodSub{\Gamma}{\gamma}{\Delta}$ can be similarly
understood in the way that $\gamma$ will output a \textbf{series} of
(dependently-typed related) terms of types $\Delta$, parametrized by
$\Gamma$ (i.e., we can consider there are holes of type $\Gamma$ in $\gamma$).
Thus (1) \ruleref{Sub Extend} looks like concatenating one
series of terms with a new term (the weird-looking type of
$\goodTerm{\Gamma}{t}{A[\gamma]}$ in the premise is due to our need of type $A$ in
the context $\Delta$ so that we can use \ruleref{Context Extension} to
create new context); (2) we have \ruleref{Proj Subst-1} to extract a
subpart of a given substitution; (3) \ruleref{Proj-Ext} explains why
substitution really looks like a (deeply-nested) pair.

Since the $\Gamma$ before the $\vdash$ is always indicating the types of
``holes'', this explains rules \ruleref{Term Subst} and \ruleref{Type
Subst}: given a term $\goodTerm{\Delta}{t}{A}$ and a substitution
$\goodSub{\Gamma}{\gamma}{\Delta}$, we will have a new term
$\goodTerm{\Gamma}{t[\gamma]}{A[\gamma]}$ after the substitution---this
substitution can be understood as ``filling all the holes''. Once all
the holes are filled with a given substitution, the new holes all come
from the substitution itself.
Consecutive substitutions can be composed using \ruleref{Sub Comp}, because
all the holes are always from the last substitutions.

>>>>

% Example : how judgemental equality is used for the universe
We review the concept of ``universe'' and at the same time showing how equality is representing semantic. In dependent type setting, function now can return type, of which the type is ``type of type'', i.e., \textit{the universe} $\cU$. We follow the style of \textit{name=ty/universe à la Tarski}~\cite{hofmann1997syntax} and
distinguish types from their \textit{codes or names} as showin in \ruleref{Code/Get Type}: given a \textit{code} (term) $\goodTerm{\Gamma}{T}{\cU}$ we can retrieve the corresponding type
$\goodType{\Gamma}{El~T}{}$, and vice-versa.\YZ{Why is it wanted to distinguish types from their codes?}

@@

\label{fig:rules:well-typed-ctx}
\judgebox{\goodCtx{\Gm}{i}}
$$ 
\Rule[name=ctx/ept]{}{\goodCtx{\cdot}{0}} 
\quad
\Rule[name=ctx/ext]
{\goodCtx{\Gm}{i} \quad \goodType{\Gm}{A}{j}}
{\goodCtx{\Gm, A}{i \cup j}}  
$$

\judgebox{ \goodType{\Gm}{T}{i} }
$$
\Rule[name=name=ty/universe]
{}
{\goodType{\Gm}{\cU}{j + 1}}
\quad 
\Rule[name=Boolean]
{}
{\goodType{\Gm}{\cB}{0}}
\quad 
\Rule[name=Bottom]
{}
{\goodType{\Gm}{\bot}{0}}
\quad 
\Rule[name=Function]
{\goodType{\Gm}{A}{j} 
  \quad \goodType{\Gm, A}{B}{k}}
{\goodType{\Gm}{\TyPi{A}{B}}{j \cup k}}
$$

@@

$$
\Rule[name=Bool]
{}
{\goodTerm{\Gm}{\true, \false}{\cB}}
\quad
\Rule[]
{\goodSub{\Gm}{\gm}{\Delta}}
{\goodType{\Gm}{ \sub{(\El{T})}{\gm} \equiv \El{(\sub{T}{\gm})} }{j} \quad
 \goodTerm{\Gm}{ \sub{\true}{\gm} \equiv \true}{\cB} \quad 
 \goodTerm{\Gm}{ \sub{\false}{\gm} \equiv \false}{\cB} 
}
$$
\judgebox{\goodSub{\Gm}{\gm}{\Delta}}
$$
\Rule[name=Ept Subst]
{}{\goodSub{\Gm}{\epsilon}{\cdot}}
\quad
\Rule[]
{}{\goodSub{\Gm}{"id"}{\Gm}}
\quad
\Rule[name=subst/comp]{
  \goodSub{\Delta}{\delta}{\Theta}
  \quad \goodSub{\Gm}{\gm}{\Delta} 
}{\goodSub{\Gm}{ \SubstComp{\delta}{\gm}}{\Theta}}
\Rule[name=subst/ext]
{\goodSub{\Gm}{\gm}{\Delta} \quad \goodTerm{\Gm}{t}{ \sub{A}{\gm}}}
{\goodSub{\Gm}{(\gm, t)}{(\Delta, A)}}
$$

@@

A term $\goodTerm{\Gamma}{t}{A}$ can be understood as one term $t$ of
type $A$, with a bunch of holes (of type $\Gamma$) inside the term $t$.
A \textbf{simultaneous substitution} $\goodSub{\Gamma}{\gamma}{\Delta}$ can be similarly
understood in the way that $\gamma$ will output a \textbf{series} of
(dependently-typed related) terms of types $\Delta$, parametrized by
$\Gamma$ (i.e., we can consider there are holes of type $\Gamma$ in $\gamma$).
Thus (1) \ruleref{subst/ext} looks like concatenating one
series of terms with a new term (the weird-looking type of
$\goodTerm{\Gamma}{t}{A[\gamma]}$ in the premise is due to our need of type $A$ in
the context $\Delta$ so that we can use \ruleref{ctx/ext} to
create new context); (2) we have \ruleref{Proj Subst-1} to extract a
subpart of a given substitution; (3) \ruleref{Proj-Ext} explains why
substitution really looks like a (deeply-nested) pair.

Since the $\Gamma$ before the $\vdash$ is always indicating the types of
``holes'', this explains rules \ruleref{Term Subst} and \ruleref{Type
Subst}: given a term $\goodTerm{\Delta}{t}{A}$ and a substitution
$\goodSub{\Gamma}{\gamma}{\Delta}$, we will have a new term
$\goodTerm{\Gamma}{t[\gamma]}{A[\gamma]}$ after the substitution---this
substitution can be understood as ``filling all the holes''. Once all
the holes are filled with a given substitution, the new holes all come
from the substitution itself.
Consecutive substitutions can be composed using \ruleref{subst/comp}, because
all the holes are always from the last substitutions.
